---
title             : "Improving the understanding of cooperation in commons through social dilemmas' literature: A scoping review"
shorttitle        : "Improving the understanding of cooperation in commons"

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "University of Bern"

authornote: |

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  According to PRISMA-ScR: Provide a structured summary that includes (as applicable) background, objectives, eligibility criteria, sources of evidence, charting methods, results, and conclusions that relate to the review questions and objectives
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "cooperation, social dilemmas, commons"
wordcount         : ""

bibliography      : CommonPaths.json

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
#bibliography: references.bib
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

PRISMA-ScR:

-   Describe the rationale for the review in the context of what is already known. Explain why the review questions/objectives lend themselves to a scoping review approach

    many studies on cooperation, but just a few focus on commons =\> similar aspects between commons and social dilemmas, so literature on social dilemma can be a powerful tool to investigate which factors could impact cooperation in commons

-   Provide an explicit statement of the questions and objectives being addressed with reference to their key elements (e.g., population or participants, concepts, and context) or other relevant key elements used to conceptualize the review questions and/or objectives.

PLAN OF THE INTRODUCTION:

-   COMMONS: what they are, why they become more and more popular and their main challenge = cooperation

    -   commons' definition

    -   more and more interest in commons because they could be a solution (see @Pacheco2014)

    -   cooperation =\> provide example of the Tragedy of the Commons

In a commons situation individual, short-term interests are at odds with collective, long-term ones [@Hardin1968]. To ensure the success of a commons, people have therefore to sacrifice a part of their personal benefits for the collective good. For instance, a fisher could be tempted to fish in large quantity from a common basin, but if every other fisher did the same, fishes would not be enough to reproduce, and therefore there would be less fish for the following season.

Cooperation, defined as a personal sacrifice for the greater good [@Kollock1998], is thus essential to in commons, and that is why a deeper understanding of how it works in this context is essential. Cooperation has often been studied with social dilemma experiments XXX.

=\> existing literature on cooperation =\> retake the text below

-   cooperation: often studied in social dilemma situation, while there is a lack of (quantitative) studies in commons

-   similarities between commons and social dilemma, that justify why taking SD to investigate commons is pertinent + to make the link: list of papers that studied a common in a social dilemma perspective (ex: Hartl & Hofmann, 2022)

**once selected the journal, re-organize all the introduction!!!**

Many societal problems that humans are facing nowadays have a social dilemma structure [@Kollock1998], i. e. they are situations in which short-term, individual interests are at odds with long-term, collective ones [@Hardin1968]. More specifically, in such situations, each individual obtains a higher personal payoff with a non-cooperative, self-interested behaviour than with a social, cooperative one, independently from what others do. However, everybody's payoff is higher if all cooperate than if all defect [@Dawes1980].

An example of real-life social dilemma is the climate crisis that we are actually facing: engaging in pro-environmental behaviours to reduce our ecological footprint is often more complicated and less advantageous for the self than acting without worrying about environment [@Griskevicius2010]. However, since behaviours that serve our personal well-being tend to be detrimental to the environment [@VanLange1998], if everyone defects by choosing the easy way, consequences will be (and already are) catastrophic. Another (less tragic) example is provided by @Kollock1998 and involves public television: the best outcome at the individual level would be to have access to the programs without paying the fee, but if nobody paid it, public television would no longer exist. Cooperation (i.e. sacrificing for the grater good at a personal cost, [@Kollock1998]) is therefore essential to avoid the worst collective outcome in social dilemma situations, and that is why a deep comprehension of this phenomenon is needed.

Classical economic theories have described human being as "homo economicus". This means that, facing a choice, individuals are considered as completely rational and always maximizing (expected) utility [@Edwards1954]. Therefore, according to these theories, individuals are not supposed to show altruistic behaviours, defined as behaviours that involve costs for the self while benefiting others [@Wittek2015]. However, there are individual differences in the level of rationality, i.e. the ability to think analytically at the situation. @Lang2018 showed that individuals scoring low in rationality cooperated more than individuals with an high score. This shows that, when investigating a phenomenon like cooperation, individual differences must be taken into account.

Another theoretical approach toward cooperation is the evolutionary perspective, who suggests that individuals can show cooperation, but only when doing so bring them a personal advantage. For example, the theory of competitive altruism [@Roberts1998] says that acting for the others' good could be a signal of a person's pro-sociality, and therefore improve this person's reputation. Moreover, the capacity to endorse a personal cost communicates strength and wealth, socially valued characteristics [@Zahavi1995]. @VanVugt2010a confirmed this assumptions in a social dilemma setting, by showing that people contributed more to a public resource when their action was visible by others, even if the public good was already provided or unattainable. In lines with hypothesis, participants who wasted their resources in the public good were perceived with an higher social status. This means that cooperation can also serve a self-interested goal, by improving the reputation of cooperators.

According to economic and evolutionary theories, individuals are always pursuing their self-interest, either by maximizing their personal payoff by defecting, either by cooperating to increase their social status and reputation. However, in social dilemma experiments, a significant amount of people are "strong reciprocators": they are willing to use their personal resources to "be kind to those who are being kind and to punish those who are being unkind", even if they are not certain that their altruistic behaviour will be rewarded [@Fehr2002; @Gachter2009]. This shows that cooperation is not always driven by personal interest and that people care about others' and environment's welfare[@Torsvik2011; @VanVugt2009]. GEF hypothesis: @VonBorgstede2018

In addiction to that, there are evidences suggesting that people do not always thoroughly analyze the context to take a rational cooperative or non-cooperative decision. In fact, cooperation seems to be the intuitive option in social dilemma experiments, while defection needs more thinking [@Kieslich2014]. Moreover, people can also cooperate in part because they are confused about the situation [@Andreoni1995]. In fact, @Fosgaard2017 showed that, in public good games, there is a considerable amount of participants that do not understand correctly the strategies to maximize the personal or the collective payoff. Moreover, these misperceptions can vary depending on how the game is presented.

Therefore, people are not always rational in the economic, individualistic way [@Kollock1998], and there are multiple factors that can influence their behaviour in social dilemmas situations [for a review, see @VanLange2013]. Understanding how these different factors impact and shape cooperation is therefore essential to try to solve many of the societal problems that we are facing. **???**

## Commons as social dilemmas

Commons are a very specific kind of social dilemma. **FIND A DEFINITION (lol) that shows the similarity with social dilemmas (common resource, ...) !!** + Since, as in social dilemmas, a conflict between individual and collective interest is present @Sussman2016 used the term of "commons dilemmas" to speak about commons.

Predictions about cooperations in commons are also similar to the social dilemma context: according to @Hardin1968 and his "Tragedy of the Commons", since individuals are likely to maximize their individual gain, commons are inevitably destined to failure due to the over exploitation of resources. However, successfully managed commons are often observed [e.g. @Ostrom1990] and, as discussed in the previous section, human beings are not motivated just by their personal interests.

@Franzen2019: example that commons can work (outlets in Swiss farms)

@Reindl2019: cooperation and defection in lab experiments are predictive of behaviours in a real public good (where individuals have to give resources to contribute to a public good)

-   But: Reindl et al. (2019) and Rommel et al. (2015)  behaviour in social dilemmas is informative about behaviours in commons + @Balliet2011: in the introduction (pages 1-2) there is a good explanation of social dilemma's utility to investigate cooperation
-   @Ros-Galvez2015: those who donated more in a game are those who post more in an online forum (= cooperation)
-   @Sussman2016: at page 11 mention that simulation of commons can be realistic why people feel real emotions (see the articles they cited) "Nevertheless, simulated commons dilemmas do reflect some degree of realism"
-   Importance of commons: Pacheco et al. (2014) having small groups behaving in an ecological way is efficient against climate change  very important to understand under which conditions cooperation could be enhanced in commons
-   @Gomez-Ruiz2020: see part about free-riders in the introduction

@Pacheco2014: "global coordination for a common good should be attempted by segmenting tasks in many small to medium sized groups, in which perception of risk is high and uncertainty in collective goals is minimized"

ideal = decentralized agreements between smaller groups, possibly focused on regional issues are prone to significantly raise the probability of success in coordinating to tame the planet's climate

-   Aim of this scoping review = shed light on cooperation both in social dilemmas and in commons by (1) identifying individual factors that can enhance it and (2) understanding what kind of organization groups can set to improve a cooperation of all members

a.  ?? Is it to mention that I focus on what a group can do, without focusing on how external institutions can influence cooperation?? If so, see @Larsson2016
b.  

# Methods

@Munn2018: reason to conduct a scoping review =\> maybe in the introduction?

To conduct this scoping review, different scientific publications were analyzed in response to the research questions. To gather the papers, a Boolean search methodology was used in different scientific search engines such as Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", "commons", "cooperation", "conflicts", "reciprocity", "trust", and they were paired with "AND" and "OR" Boolean operators. The research was made during the month of August, 2023. According with @Munn2022, the type of source was not a selection criteria, therefore journal articles, conference papers and book chapter could be included. This allows us to have an idea as complete as possible about the existing literature on cooperation in commons and social dilemmas. Moreover, there was any chronological or geographical limitation, once again to provide a broader vision of the topic. There were however some selection criteria concerning the papers' methodology and content. First of all, as mentioned in the introduction, this review will focus on individual and group dynamic that can enhance or decrease cooperation in a social dilemma situation (i. e. commons). For this reason, all the papers investigating how an external entity/institution can shape cooperation were not included. Secondly, papers that studied social dilemmas with an agent-based model methodology were also excluded. **include explication** Moreover, in case of experimental or correlative studies, only cases with adult participants were selected.

Initially, 5210 papers emerged from the key phrases search. After deleting the duplicates, the count was of 4586. Based on titles and abstracts, 4321 articles were eliminated, leaving 253 texts to examine. After a full-text evaluation of the remaining papers, 91 were selected for the present review. Figure 1 resumes this selection procedure.

## Papers' description

The 91 included papers were published between 1984 and 2023, with a median of 2015. Almost all of the selected texts were journal articles, with the only exception being a book chapter. Most of them were experimental studies (*N*=62), followed by literature review and meta-analysis (*N*=14) and correlational studies[^1] (*N*=14). The remaining paper was a qualitative study.

[^1]: With "experimental studies" are intended all the studies were at least a variable is manipulated by researchers, while "correlational studies" refers to papers using quantitative analysis, but without manipulations.

The papers' topics can be divided in 4 categories: first, "social dilemma setting" (*N*=77) refers to high controlled laboratory experiments where participants have to take cooperative decision. Usually some variables are manipulated to investigate the influence of a specific factor on cooperation [e.g. @Zhang2019]. The type of paradigm used was most of the time a public good game (where participants can decide how much to contribute to a public good, e.g. @Torsvik2011) or a common resource game (where participants can decide how much to take from a common resource, e.g. @Nockur2020). Moreover, the type of game could change vary during the experiment [e.g. @Fosgaard2019]. The second category is "real-life social dilemma situation" (*N*=7), and investigate cooperation in different real-life situations that have a social dilemma structure such as recycling behaviours [@Rompf2017] or use of public transportation [@VanLange1998]. Thirdly, "real-life commons" (*N*=4) refers to studies that investigate cooperation within participants that are members of a commons in their everyday life. Finally, papers in the category "commons modeled by a social dilemma experiment" (*N*=3) use a laboratory social dilemma paradigm, but with clear references to a concrete common. For instance, in @Mosler1993, participants had to imagine to be fishermen who shared a common basin.

Most of the papers concerned psychology and economic. In fact, thirty-five papers were published in psychology journals, such as *Group Processes & Integroup Relations* or *Psychological Bulletin*, and fourteen in economic journals (e.g. *Experimental Economics*). Additionally, some papers (*N*=8) were published in interdisciplinary journals such as *Journal of Economic Psychology.*

Finally, most of the studies were conducted in Europe (*N*=40) (especially in Germany (*N*=12) and in the Netherlands (*N*=9)), followed by North America (*N*=20, of which *N*=19 in the USA) and Asia (*N*=9, of which *N*=6 in China). Moreover, *N*=7 studies were set in more than one country, and *N*=13 were meta-analysis or literature reviews. The two remaining studies took place in Australia and in Colombia, respectively.

# Individual factors

## Gender

The effect of gender on cooperation is controversial, and studies that investigated it provided contradictory results: sometimes men cooperated more [@Dorrough2019], sometimes women did [@Peshkovskaya2017], and sometimes there was no difference between genders. In general, meta-analysis found a non significant effect size for this difference [e.g. @Balliet2011; @Spadaro2023]. This indicates that gender differences in cooperation are very flexible and can vary depending on the context [@Spadaro2023]. Therefore, when a significant effect is observed, it means that gender has somewhat been made salient by the situation [@Sell2009].

One of the most observed moderator of this effect is the interaction partners. In fact, in a social dilemma experiment, women were more cooperative toward strangers than men, while men cooperated more in a group of friends. Moreover, after a socialization phase where strangers could interact, men cooperated more than women [@Peshkovskaya2017]. These results seem to indicate that men accord more importance at the relationship with other members than women. @Gomez-Ruiz2020 also got to the same conclusion by showing that the creation of a team identity enhanced men's cooperation, but had no effects on women. Moreover, in a social dilemma with many iterations, men, but not women, tended to increase their cooperation over time [@Balliet2011].

The partners' gender also have an impact on women's and men's cooperation. In general, groups composed only by men are more cooperative than only-women groups. This is explained by the fact that women are less achievement oriented and more socially oriented than men: they are therefore less concerned by the maximization of their personal payoff and more about relative payoff. Thus, the defection strategy allows them to play defensively and to avoid earning less than the other players [@Colman2018]. Moreover, men displayed higher levels of gratitude and trust in same-sex groups than women [@Peshkovskaya2019]. On the other side, women cooperate more in mixed group. In fact, when both men and women are present, gender stereotypes are activated and people tend to adapt their behaviour to them. This leads women to behave in a more prosocial, altruistic way and men to be more self-oriented and competitive [@Balliet2011]. The same result is also found by @Peshkovskaya2019, who also demonstrated that women reported a stronger social connection with their partners in mixed than in same-sex groups. The overall group composition also have an impact on cooperation: in general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023]. However, @Barrero-Amortegui2021 showed that the level of resource exploitation is linked to the number of women in a group, but that this link is mediated by the signature of an agreement about extraction. More precisely, in their experiment, groups with a majority of women extracted more resources from a common good (and thus cooperated less) before the signature of the agreement, but also showed the largest reduction of the exploitation level after the agreement was signed. Therefore, even the effect of group composition seems to be subjected to moderator variables.

Another reason why gender differences in cooperation are often very small is that fear and greed motives are confused in a classical social dilemma situation [@Simpson2003]. In this contexts, a feeling of fear arises when one's cooperation could be exploited by others' free-riding, while greed is the feeling associated with the temptation to free-ride on others' cooperation. According to social psychology and evolutionary theories, the author expected women to defect more when fear incentives were present, and men to defect more with greed incentives. Although he could corroborate only the second hypothesis, his intuition about women being more sensitive to the fear of being exploited is confirmed by other studies. For example, women level of cooperation is predicted by their trust in their interaction partner, while men's trust and cooperation are not linked [@Irwin2015]. Moreover, women's lower cooperation is explained by the fact that they expected their partner to contribute less to the common good [@Dorrough2019].

Finally, some of the analyzed papers presented interesting secondary results. First, against their authors' expectations, some intercultural studies showed that the gender difference in cooperation is not explained by the level of gender equality of the country [@Dorrough2019; @Spadaro2023]. Secondly, in an experiment not involving a social dilemma, women were significantly more altruistic than men [@Dorrough2019]. According to the authors, this underlines the importance of distinguishing between prosociality and strategic interactions that characterize social dilemmas. Therefore, in a situation where classical attributes of social dilemmas are not explicit, gender difference in cooperation could not be subjected to the same moderator effects described above. Lastly, according to @Sell2009, it is not possible to think about gender without relying it to social status, with men representing the high-status group. Therefore, gender differences in cooperation could also be caused by social status: to have a better understanding of this hypothesis, it would be interesting to investigate the difference in cooperation between two other social groups with a different status (e.g. high and low socioeconomic status).

In conclusion, "at times women are more cooperative than men, less cooperative than men, or equally as cooperative as men. These variations are explained, in large part, by structural differences and identities triggered by those differences" [@Sell2009]. The context in which men and women interact is therefore essential to predict if a gender will cooperate more than the other, and which one.

## Social status

Few studies have investigate how social status impact cooperation, and their results do not allow to draw a clear conclusion about this effect.

For instance, advantaged members (i.e. high-status class) tended to cooperate less than disadvantaged ones [@Aksoy2019]. In fact, when participants did not receive the same endowment at the beginning of the experiment, contributions of "rich" and "poor" members converged [@Kingsley2016]. In other words, proportionally to their resources, low-status participants cooperated more. However, in their review, @VanLange2013 identified some studies that observed the opposite effect, namely that wealthier people (or those who benefitted more from a common resource) contributed more to the common good than the other.

Social status can also create inter-class conflicts. In fact, people cooperated more with the members of the same social class than with other classes' members [@Aksoy2019]. Moreover, a form of solidarity is observed between low-status members: in a study where participants had the opportunity to punish others for a non-cooperative behaviour, low-status members punished significantly less other low-status participants [@Chen2022]. Also, in an experiment where participants were randomly advantaged during the interaction, @Camera2020 found that subjects who received a lot of advantages were discriminated by others, and especially by the ones who were rarely advantaged.

Regarding the overall group composition, results are not consensual. On one hand, @Aksoy2019 found that groups with an inequality of status between members displayed lower levels of cooperation than homogeneous groups. On the other hand, economic and socio-cultural heterogeneity, although being negatively correlated with in-group trust, does not have a direct effect on a common-pool resource well-being [@VanKlingeren2021].

In conclusion, social status impacts cooperation, but it seems to have multiple effects. Moreover, as mentioned in the previous section, @Sell2009 affirmed that, when investigating gender differences in cooperation, it is impossible to disentangle the effect of social status. Similarly, @Aksoy2019 demonstrated that the lower level of cooperation between members of different social classes is mainly due to actors' identification with their classes. This shows that social identity and social status are deeply connected, and that therefore it is difficult to separate their effects on cooperation.

## Group identity

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group [e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020]. For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Group identification can also esaily be implemented in a laboratory experiment, and it has still a strong positive effect on cooperation [@Wit1992]. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high identified individuals cooperated more than low identified ones when faced to an in-group participant who defected (and this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption, while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

However, despite the observed beneficial effects of inducing an intergroup competition, in a real-life situation this could be harmful. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by the competitive context, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

Finally, social identity does not only have a main effect on people's cooperation, but can also moderate other factors' impact. For example, monetary incentive are more effective at enhancing cooperation of low identified individual than of high identified ones [@VanVugt2009]. Moreover, a previous group failure can decrease cooperation between low identified member, but this effect is not observed in people who have a strong identification with their group [@Jackson2011; @Jackson2012].

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt2009]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful technique to enhance cooperation in an easy and cheap way.

## Values and personality traits

Personal values and personality traits can play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation is very consensual: in a meta-analysis, 81 out of the 82 papers examined reported than pro-social cooperated significantly more that pro-self [@Balliet2009].

However, this general effect is subjected to several moderators. For example, pro-self individuals cooperated more in a game where economic incentives to cooperation were present than in a game without such incentives, while pro-social did not displayed this difference and showed high levels of cooperation in both contexts [@Emonds2011]. A similar effect is also described by @Balliet2009, who found a stronger effect of SVO in unpaid games than in paid ones, because, when monetary rewards were present, pro-self increased their cooperation. Therefore, pro-self individuals also cooperate, but only when cooperation serves their personal interest. In fact, in a study about commuting habits, the extent to which the cooperative behaviour (i.e. choosing to commute by public transportation or carpooling) was perceived as efficient impacted pro-self individual's decision more than pro-social's [@VanLange1998].

Moreover, SVO has a stronger effect if there is an uncertainty about the number of individuals having access to a common resource: when the number of other participants was not known, pro-social, but not pro-self, reduced their harvesting comparing to a group-size certainty condition [@DeKwaadsteniet2008]. The same authors demonstrated that this result is explained by the fact that pro-social individuals want to maximize the joint outcome, while pro-self only think at their own personal interest. Also, SVO effect is partly explained by trust, i.e. the extent to which people think that their interaction partners will also cooperate. In fact, in @VanLange1998, pro-social individuals with a low level of trust did not make different decisions than pro-self individuals, while the most cooperative were pro-social with an high level of trust. Finally, @Lu2019 showed that the intuitive choice of pro-social individuals is cooperation, while for pro-self is defection. However, pro-self can cooperate more if they have the possibility to deliberate about the situation, because they can realize that the best strategy to maximize the joint (and thus personal) outcome is cooperation.

Other studies also focused on how much people are self- or social-oriented by using other constructs. For example, @Liu2009a demonstrated that priming a common fate to induce a context of interdependence significantly increased the cooperation level of participants with an independent self, but had no effect on participants with an interdependent self (who already displayed high levels of cooperation). Moreover, in a social dilemma experiment, competitive, individualistic people were motivated just by personal gain, while people who saw themselves as a part of a group were more interested on maximizing the income of each member, even at the cost to have a smaller personal payoff [@Probst1999]. Also, high allocentric (i.e collectivist) individuals cooperated more than low allocentric ones when the perceived social norms are cooperative: this means that they are more sensitive to social cues and therefore they adapt their behaviour to social norms [@Chen2007].

Finally, there are other personal values and traits that can affect cooperation. For example, pro-environmental values are negatively correlated with harvesting from a common resource, but only when the resource is scarce; therefore these values are more likely to motivate difficult, rather than easy, ecological behaviours [@Sussman2016]. Levels of harvesting from a common resource are also lower between left-wingers than right-wingers [@Fosgaard2019]. Moreover, humility and honesty are significant predictor of cooperation [@Kieslich2014], and this is due to the fact that people with such characteristics manage to resist to the temptation to exploit others' contributions [@Hilbig2018]. Self-control also plays a role, in the sense that more self-control leads to a greater cooperation [@Kocher2017]. Lastly, people with an higher justice sensitivity make bigger contributions to a public good [@Schlosser2018].

In conclusion, values and personality traits have a large impact on human cooperation, and it is thus very important to take them into account when investigating this phenomenon. Interestingly, some social psychological techniques (e.g. elaboration) allow to modify an individual's values [@Sussman2016]. Such techniques could therefore be used to improve cooperation.

## Trust and beliefs about others' contribution

In social dilemma situations, trust can be defined as the "expectations of others' benevolent motives in situations that involve a conflict between self and collective interests" [@Balliet2013, p.1]. Trust can be dispositional or a state; the former is an individual's general trait, and is usually measured with personality scales, the latter is linked to the context, and can be operazionalized by expectations about a partner's behaviour in a specific situation [@Balliet2013]. The two are positevely correlated [@Lubke2021], and therefore both have a positive general effect on cooperation [@VanLange2013; @Balliet2013]. In fact, high levels of trust are associated with more contributions to a public good [@Kocher2015], and a participant's contribution in a public good game can be predicted by her beliefs about others' contributions [@Smith2013]. These beliefs are linked with SVO orientation [@Ackermann2019].

Several studied confirmed this relationship. For instance, @Chaudhuri2002 demonstrated that people who cooperated more in a first public good game trusted more their partner in a following interaction than people whose main strategy was defecting. Moreover, people that were pessimistic about others' contribution cooperated significantly less than the optimistic [@Oyediran2018], and those who free-rode in a public good game, compared to cooperators, believed that others contributed less [@Kocher2015]. Also, people were more willing to purchase actions for a CO2 mitigation initiative when they believed that the majority of others had also bought them [@Sturm2019]. Finally, in a study about self-administered outlets in Swiss farms (i.e. outlets where nobody monitors if people pay or not their purchase), @Franzen2019 found that a farmer's general trust in other people is a significant predictor of the probability of having such an outlet. Concerning the moderators, trust matters more in situations involving a large, rather than small, conflict between personal and collective interests [@Balliet2013], and when people are uncertain about others' intentions [@VanLange2013].

People can also display different levels of trust in the "reliability, effectiveness and legitimacy of public institutions" [@Rompf2017, p.2]. This authors showed that institutional trust is associated with the probability of recycling, because knowing that sanctions can be efficiently applied can motivate people to adapt to norms. Moreover, the effects of perceived costs and benefits of the action are less predictive in people with high institutional trust.

Finally, trusting others (and the institutions) is essential to display higher level of cooperation. However, general beliefs about others' cooperation are quite low compared to the actual behaviours, and this trust gap can undermine cooperation, since people's actions are significantly predicted by their beliefs [@Lubke2021]. Therefore it is essential to provide people with reliable information about others' behaviour, to increase trust and, as a consequence, cooperation.

# Group factors

## Incentives

Cooperation rates can be modified by introducing incentives, who can be either punishments or rewards[^2]. In the litterature, punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2015], and its positive effect on cooperation is very consensual. In general, when there is the possibility for participants to punish others members based on their contributions, cooperation is higher than when this is not possible [e.g. @Pfattheicher2018; @Balliet2011a]. Moreover, the presence of a punishment system increases the perception of fairness [@Nockur2021].

[^2]: Since the present paper's interest is to investigate how the internal organization of a group can affect cooperation, only papers that studied situations where the participants could give rewards or punishments to each-others are included. Incentives given by third party have also an effect on cooperation, but they are not included in this review.

However, this main results is subjected to several moderator factors. First, the punishment have a bigger impact on cooperation if its implementation is costly for the punishers than if it is free [@Balliet2011a], but at the same time the cost should not be too high [@Egas2008]. Moreover, punishments are more effective if the interaction is among friends than among strangers, and if people know that they will have to interact repeatedly with each other [@Balliet2011a]. In fact, in short interactions, if punishment is costly, individuals who punish others are the ones with the lowest personal income [@Milinski2012], and this could discourage the application of the fine. Also, an experiment showed that a punishment system increased cooperation in groups where all members had the same amount of resources, but not in groups where income was heterogeneous [@Kingsley2016]. In fact, @Chen2022 found that, in groups where the initial endowment was heterogeneous, participants that received less didn't punish the others "low-status" members. This solidarity could therefore encourage free-riding. Moreover, the effect of punishment is stronger if only cooperators have the power to punish defectors, than if everybody can punish [@Xiao2016]. Finally, a plain situation is also needed to have a stronger effect of punishment: the link between behaviour and outcome must be clear, without random factors affecting the final outcome [@Xiao2016], and the punishment must be highly effective [@Egas2008].

However, punishment must be applied with carefulness: @Rockenbach2019 compared the punishment strategy of different groups, and found that the most successful ones (whose members had the highest payoff at the end of the experiment) punished with caution and not in early stages of the game. In these groups, a norm of when and how punishment should be applied seems to exist. Moreover, punishment can also have two main negative sides. First, it could entails a second-order social dilemma [@Milinski2012], since cooperative participants can free-ride on the effort of other cooperators who punish non-cooperators. Secondly, anti-social punishment can arise. This term refer to the punishment of cooperators by non-cooperators, and it is present when all participants can punish independently of their contributions [@Xiao2016]. In this context, defectors can see cooperators as rule breakers, and anti-social punishment is a sort of retaliation [@VanDijk2015].

A possible solution could be to implement a system of democratic punishment, where a fine is given only if a majority of participants agree. In fact, compared to a peer punishment system where participants decide individually if and who punish, in a democratic system anti-social punishment is less present, general punishment level is lower, and final payoff and cooperation rates are higher [@Pfattheicher2018].

Rewarding systems, on the other side, are less investigated. Globally, they have a positive effect on cooperation [@Balliet2011a]. However, some studies found that they were more efficient than punishment [e.g. @Chen2022], while others came to the opposite conclusion [e.g. @Milinski2012] or found no difference between the two systems [e.g. @Balliet2011a].

In a meta-analysis, @Balliet2011a found that rewards had a stronger effect on cooperation if their implementation was costly rather than free, and in iterated, rather than single, interactions. Moreover, rewarding was more effective in situations where participants had to decide how much take from a common resource, compared to when they had to decide how much contribute to a common good.

In general, incentives seem to be a powerful tool to enhance cooperation, but they have a dark side. In fact, they reduce participants' intrinsic motivation [@VanDijk2015], by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. In fact, @Schlosser2018 demonstrated that people's level of justice sensitivity could predict their contributions to a public good game, but this link disappeared when a punishment system was introduced. Also, when incentives are present people think that others cooperate just because of them, and this undermines their trust in other members[@Irwin2014], and especially the trust in the fact that others are internally motivated to cooperate [@Mulder2006]. Therefore, if incentives are removed, cooperation drops [@Irwin2014; @Mulder2006].

Finally, incentives are often studied in economic formats, but they can also be social. For example, the expectation of honour or shame (after cooperating or defecting, respectively) is enough to enhance cooperation [@Jacquet2011]. Moreover, expressing disapproval toward defectors can boost cooperation by inducing more compliance to the cooperative social norm [@Nelissen2013]. Even an ostracism system where participants can vote to exclude other members from the game is very effective in increasing cooperation [@Maier-Rigaud2010]. In general, social sanctions can be even more efficient than economic ones, because cooperation do not drop after their removal [@Nelissen2013].

To conclude, economic incentives can have a positive effect on cooperation, but their efficient implementation must be insured all along the interaction, otherwise cooperation rates would drop. On the other hand, social incentives seem to have the same positive effect, and their removal do not cause a reduction of cooperation. However, there is a lack of studies focusing on this kind of sanctions.

## Communication

Communication has a global positive effect on cooperation [e.g. @Balliet2010; @Bicchieri2002], but this effect is subjected to several moderators. First, talking is more powerful than sending written messages, because it allows non-verbal communication and facilitates the social norms formation process. Moreover, communication's effect is stronger in large groups: when many people are present cooperation rates are lower, so communication between members can buffer this negative effect [@Balliet2010]. Finally, if communicate with others requires a personal cost, the effect is weaker than if communication is free [@Janssen2014].

The content of the messages is also essential. @Koessler2021 identified four steps that need to be present when participants in a social dilemma situation communicate: problem awareness, identification of strategies, agreement and ratification. In another study, groups who used communication to foster a common social identity, promote cooperative social norms and ensure that all the players publicly committed displayed higher rates of cooperation than other groups [@Adams2022]. However, casual chit-chat also plays a very important role in increasing cooperation [@Koessler2021]. Also, the increase of cooperation through communication can be explained by the normative content that people are shaping while exchanging messages [@Janssen2014], and by the fact that it allows participant to reach a consensus about the best strategy [@Hopthrow2005]. Moreover, even if communication also helps to create a social identity, the effect of social norms is a stronger mediator of the link between communication and cooperation [@Bicchieri2002].

Another factor that influences the effect that communication has on cooperation is whether individuals exchange messages before, during or after the cooperative interaction phase. Concerning the post-game communication, @Kumakawa2013 found that receiving a feedback about the cooperation choice made previously can increase cooperation in the following interactions, especially if the feedback is negative. However, just the fact of knowing that a discussion will follow is not enough to enhance cooperation, and can even have a counterproductive effect, because forcing people to think about a future discussion make them reflect more on the situation and thus act more rationally [@Torsvik2011]. On the other hand, communication before or during the interaction have both a positive impact on cooperation, and the difference between their effect is not significant [@Balliet2010].

Finally, several studies investigated the effect of gossip as a form of communication that can increase cooperation. Gossip is defined as a spread of reputational information about another person [@Feinberg2014], and it increases cooperation in two ways: first, if people know that gossip is possible, they tend to cooperate more to avoid the risk of being the target of such gossip and therefore of being excluded from following interactions. Secondly, having information about the past behaviour of potential cooperation partners can help to select the most cooperative people in an interaction situation [@Feinberg2014]. Moreover, the same authors found that pro-social individuals gossiped more than pro-self. This shows that gossip can have the pro-social function of protecting others from a potential exploitation [@Feinberg2012]. However, gossip has also some self-interest effects: for instance, observing another subject being exploited and having the opportunity to gossip about the exploiter significantly reduced negative affects in @Feinberg2012 's participants. Moreover, providing correct information while gossiping helps to build a positive reputation [@Giardini2021].

In conclusion, when people have the possibility to communicate before, during or after a cooperative interaction, the cooperation level increases. Even if the strength of this effect can vary depending on the situation, communication never seems to have a negative impact on cooperation.

## Social Norms

Social norms can provide a guide on how to behave, especially when there is uncertainty about others' behaviour [@VonBorgstede2018]. @Thogersen2008 studied the effect of social norms on daily social dilemma situations such as environmental behaviours, and found that both injunctive (what others approve) and descriptive (what others do) are positively related to behaviours. Moreover, cooperation can be maintained via social norms, because people who do not respect them are susceptible of social sanctions. In fact, if social approval is possible, people cooperate more [@Rege2004].

There are several explanations of social norms' effect on cooperation. Firstly, according to the appropriateness framework, when taking a decision people ask themselves: "What does a person like me do in a situation like this?" [@Arora2012]. Therefore they need to define the situation as a cooperative task and to strongly identify with the group [@VanLange2013]. This leads to a greater accessibility of group's norms and consequently (if the norms are cooperative) to more cooperation. Testing this assumption, @Arora2012 showed that a greater identification with social group lead to higher levels of cooperation. The second explanation concerns conformity: in general people adapt their cooperation to others' behaviour, but this is usually explained by the reciprocity principle (cooperate enhances the probability that others will do the same, and therefore that the final payoff will be higher) [@Bardsley2005a]. However, these authors demonstrated that people conform to others' behaviour even if this has no impact on their personal payoff. Therefore, people have a general tendency to imitate others to follow social norms.

Finally, social norms can be modified by participants. For instance, @Weber2008 demonstrated that in several social dilemma experiment there are consistent cooperators, i.e. people who always cooperate. The presence of these people can increase other members' cooperation, and this effect is mediated by the perception of the social norms. Moreover, social norms can also be shaped via communication [@Bicchieri2002].

In conclusion, social norms have a significant effect on people's cooperation in social dilemma situations. Since they are quite easily modified [@Weber2008], shaping cooperative norms can be a powerful instrument to improve cooperation.

## Anonymity

There are not many studies on the influence of anonymity on cooperation, but, according to @Wang2017, most of the factors that can increase cooperation rely on a reduction of anonymity. For instance, it is not possible to punish defectors without knowing who they are. Therefore, in general it seems that non-anonymous situations can significantly increase cooperation rates. In fact, when anonymity is removed in a social dilemma situation, cooperation increases and defection is reduced [@Wang2017]. This can be explained by the fact that, when their actions are visible, people worry about others' approval [@Rege2004] and that, in non-anonymous situations, contributions can be a self-presentation strategy, because being perceived as pro-social can increase one's social status [@VanVugt2010a]. There are therefore more social incentives to cooperate in non-anonymous, rather than anonymous, situations.

Moreover, this effect is displayed especially in situations when individuals face a cost when cooperating and know that they will not get any personal benefit. For instance, @Butz2021 demonstrated that, in a public good game where a part of contributions was donated to a charity association, disclosing identity increased cooperation, but only if the donation was deducted from participants' payoff. In the same direction, @VanVugt2010a found that, when actions were public, contributions were higher if the public good was already attained or unattainable, and therefore if participants did not have a personal interest in cooperating.

Reducing anonymity in cooperative interactions can be very effective to encourage people to contribute more. A possible way to do this is to reduce group's size. In fact, according to @Romano2016, a larger group leads to de-individuation, because it is more difficult to understand who is defecting. Therefore, it is easier for members to feel anonymous and to perceive the social punishments associated with defection as less probable. Moreover, communicate is easier in smaller groups [@VonBorgstede2018]. This assumption is confirmed experimentally by @Jiang2021, who showed than in smaller groups people contributed more often to a public good. Increasing cooperation through the reduction of anonymity is also possible by giving public feedbacks about people's cooperation [@Przepiorka2020].

Another effective method to reduce anonymity is to make people publicly promise to cooperate before the cooperative interaction. For instance, in a laboratory experiment modeling a common pool resource, people who publicly committed exploited significantly less the resource than people who committed in private, even if after committing they were not restricted in their harvesting possibilities [@Mosler1993]. This results was also confirmed by @Przepiorka2020, who showed that non-binding promises increased cooperation. Moreover, even a compulsory commitment has a beneficial effect on cooperation, even if the most efficient groups are those where all participants signed voluntary the pledge [@Koessler2021a].

In conclusion, reducing anonymity during interactions is essential to improve cooperation, because this can facilitate the implementation of others cooperation-enhancing techniques [@Wang2017]. Moreover, it exists several low-cost method to make situations non-anonymous.

# Discussion

individual factor: gender is interpretable by social status and ss is interpretable by group identity

group factors: communication linked to social norms and social identity + anonymity underlies all the others + incentives has dark sides

## Limitation

observation: majority of papers included are about SD, and very few studies about commons are quantitative --\> future research: quantitative approach to investigate cooperation in common initiatives

critiques at SD because it is artificial, does not take into account all real-life factors --\> so it is good to keep it as a base, but investigations in commons are needed

60/91 papers in Europe and North America + reviews also based on studies in these countries =\> bias (talk about cultural differences in cooperation)

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
