---
title             : "Cooperation in commons and social dilemmas: A scoping review"
shorttitle        : "Cooperation in commons and social dilemmas"

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "University of Bern"

authornote: |

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "cooperation, social dilemmas, commons"
wordcount         : ""

bibliography      : CommonPaths.json

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
#bibliography: references.bib
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Many societal problems that humans are facing nowadays are social dilemmas, i. e. situations in which short-term individual and long-term collective interests are at odds [@Hardin1968]. More specifically, independently of others' actions, every individual obtains a better individual payoff by acting in an egoistic, non-cooperative way. However, if everybody did so, the worst possible outcome is obtained ????

where each individual's payoff is higher with an egoistic, non-cooperative behaviour (no matter what others do), but where the best individual outcome is reached when everybody cooperates [@Dawes1980].

Reaching a better understanding of mechanism underlying cooperation in social dilemmas is even more important when considering that many everyday situations have a social dilemma structure. An example is the climate crisis that we are actually facing: engaging in pro-environmental behaviours to reduce our ecological footprint is often more complicated and less advantageous than acting without worrying about environment [@Griskevicius2010]. However, if everyone defects by choosing the easy way, consequences will be (and already are) catastrophic. Another (less tragic) example is provided by @Kollock1998 and involves public television: the best outcome at the individual level would be to have access to the programs without paying the fee, but if nobody pays it, then public television will stop existing. + @Kollock1998: page 183, many of the problems that we face are, at their core, social dilemmas. A third example of real-life social dilemmas are commons. (**USE THIS AT THE VERY BEGINNING TO INTRODUCE THE SOCIETAL PROBLEM?**)


introduction of the societal problem = cooperation (= renonce to the individual most advantageous choice for the common good) is the key to resolve many societal problems (e.g. covid vaccination, climate change)

Classical economic theories have described human being as "homo economicus". This means that, facing a choice, individuals are considered as completely rational and always maximizing (expected) utility [@Edwards1954]. Therefore, according to these theories, human beings are not supposed to show altruistic behaviours, i.e. behaviours that involve costs for the self while benefiting others [@Wittek2015]. However, prosocial acts are often observed in human interactions (**FIND REFERENCES**), and many theories have been developed to explain them. An example is the theory of competitive altruism [@Roberts1998], that says that acting for the good of others could be a signal of a person's prosociality and therefore improve this person's reputation. Moreover, in the evolutionary perspective, the capacity to endorse a personal cost communicates strength and wealth, socially valued characteristics [@Zahavi1995]. Adding something on cooperation norms? Maybe this could be the link between altruism and cooperation? See also the article in Zotero that says that altruism is one of the seven forms that cooperation could have, so maybe is it better to talk about cooperation from the beginning???

@Torsvik2011: in the introduction says that people in social dilemma are confused and can make random choices = argument against the completely rational human (see wich papers they cited) <doi:10.1098/rstb.2010.0135>: in the introduction they cite papers that have observed cooperation

**??? CONNECTING ALTRUISM WITH COOPERATION ???** Cooperation = sacrificing for the grater good at a personal cost (Kollock, 1998)  **\<- EXPLAIN BETTER!!!**

-   @Kramer1984: see the phrases underlined in the introduction

Thus, the most rational behaviour for everyone would be to defect, because it would ensure the maximum outcome independently of other individuals' choices. ADD NASH EQUILIBRIUM??? (@Colman2018 = at page 2 good definition of Nash) However, as with altruism, cooperative decisions are often made in social dilemmas, showing that human beings are not always rational in the economic, individual sense [@Kollock1998] and that there are multiple factors that can influence people's decisions in such situations (see for example @VanLange2013 for a review).

b.  Moreover, some studies consider cooperation to be the intuitive option, while defection needs more thinking (Kieslich & Hilbig, 2014)
c.  A deeper understanding of mechanisms underlying cooperation in social dilemma is therefore necessary

-   cooperation even more fragile when scarce resources (De Luca et al.) =\> importance of understanding it


-   @VanLange1998: many behaviours that tend to serve our personal well-being are detrimental to the environment

@Jackson2012: according to economist, cooperation is an irrational response in social dilemmas (see the articles that he cites)

-   This is even more important if we consider the practical implication that social dilemmas have in our society: social dilemmas could be translated into commons

a.  Climate change as social dilemmas
b.  Definition of commons  commons as social dilemmas
c.  Article that says that social dilemmas are not the same as commons (Cole & Grossman, 2010)
d.  But: Reindl et al. (2019) and Rommel et al. (2015)  behaviour in social dilemmas is informative about behaviours in commons + @Balliet2011: in the introduction (pages 1-2) there is a good explanation of social dilemma's utility to investigate cooperation

-   @Sussman2016: at page 11 mention that simulation of commons can be realistic why people feel real emotions (see the articles they cited)

e.  Existing theory about the commons = similar structure than for social dilemmas: rational human beings are supposed to fail (TOC, Hardin, 1986), but in reality cooperation is often observed (for example Ostrom, 1990 + find others)

-   Importance of commons: Pacheco et al. (2014) having small groups behaving in an ecological way is efficient against climate change  very important to understand under which conditions cooperation could be enhanced in commons
-   Aim of this scoping review = shed light on cooperation both in social dilemmas and in commons by (1) identifying individual factors that can enhance it and (2) understanding what kind of group organization groups can set to improve a cooperation of all members

a.  ?? Is it to mention that I focus on what a group can do, without focusing on how external institutions can influence cooperation?? If so, see @Larsson2016

-   Research questions (see if not too redundant with the previous phrase)

-   introducing free-riding and defection concept somewhere (see @Gomez-Ruiz2020 introduction for free-riding)

# Methods

To conduct this scoping review, different papers published in academic journals were analyzed in response to the research questions. To gather the papers, a Boolean search methodology was used in different scientific search engines such as Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", "commons", "cooperation", "conflicts", "reciprocity", "trust" **, "intergroup"**, and they were paired with "AND" and "OR" Boolean operators. The research was made during the month of August, 2023. According with @Munn2022, the type of source was not a selection criteria, therefore experimental and qualitative studies, as well as meta-analysis and systematic reviews are included. This allows us to have an idea as complete as possible about the existing literature on cooperation in commons and social dilemmas. Moreover, there were not chronological or geographical limitations, once again to provide a broader vision of the topic. There were however some selection criteria concerning the papers' content. First of all, as mentioned in the introduction, this review will focus on individual and group dynamic that can enhance or decrease cooperation in a social dilemma situation (i. e. commons). For this reason, all the papers investigating how an external entity/institution can shape cooperation were not included. Secondly, papers that studied social dilemmas with an agent-based model methodology were also excluded. **include explication**

Criteria developed during screening: - no papers about agent-based model - no ethical/religious/philosophical papers - adult participants

Initially, 5210 papers emerged from the key phrases search. After deleting the duplicates, the count was of 4586. Based on titles and abstracts, 4321 articles were eliminated, leaving 265 (**with emotions**) texts to examine. After a full-text evaluation of the remaining papers, XXX were selected for the present review. Figure 1 resumes this selection procedure.

## Papers' description

# Individual factors

## Gender

The effect of gender on cooperation is controversial, and studies that investigated it provided contradictory results: sometimes men cooperated more [@Dorrough2019], sometimes women did [@Peshkovskaya2017], and sometimes there was no difference between genders. In general, meta-analysis found a non significant effect size for this difference [e.g. @Balliet2011; @Spadaro2023]. This indicates that gender differences in cooperation are very flexible and can vary depending on the context [@Spadaro2023]. Therefore, when a significant effect is observed, it means that gender has somewhat been made salient by the situation [@Sell2009].

One of the most observed moderator of this effect is the interaction partners. In fact, in a social dilemma experiment, women were more cooperative toward strangers than men, while men cooperated more in a group of friends. Moreover, after a socialization phase where strangers could interact, men cooperated more than women [@Peshkovskaya2017]. These results seem to indicate that men accord more importance at the relationship with other members than women. @Gomez-Ruiz2020 also got to the same conclusion by showing that the creation of a team identity enhanced men's cooperation, but had no effects on women. Moreover, in a social dilemma with many iterations, men, but not women, tended to increase their cooperation over time [@Balliet2011].

The partners' gender also have an impact on women's and men's cooperation. In general, groups composed only by men are more cooperative than only-women groups. This is explained by the fact that women are less achievement oriented and more socially oriented than men: they are therefore less concerned by the maximization of their personal payoff and more about relative payoff. Thus, the defection strategy allows them to play defensively and to avoid earning less than the other players [@Colman2018]. Moreover, men displayed higher levels of gratitude and trust in same-sex groups than women [@Peshkovskaya2019]. On the other side, women cooperate more in mixed group. In fact, when both men and women are present, gender stereotypes are activated and people tend to adapt their behaviour to them. This leads women to behave in a more prosocial, altruistic way and men to be more self-oriented and competitive [@Balliet2011]. The same result is also found by @Peshkovskaya2019, who also demonstrated that women reported a stronger social connection with their partners in mixed than in same-sex groups. The overall group composition also have an impact on cooperation: in general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023]. However, @Barrero-Amortegui2021 showed that the level of resource exploitation is linked to the number of women in a group, but that this link is mediated by the signature of an agreement about extraction. More precisely, in their experiment, groups with a majority of women extracted more resources from a common good (and thus cooperated less) before the signature of the agreement, but also showed the largest reduction of the exploitation level after the agreement was signed. Therefore, even the effect of group composition seems to be subjected to moderator variables.

Another reason why gender differences in cooperation are often very small is that fear and greed motives are confused in a classical social dilemma situation [@Simpson2003]. In this contexts, a feeling of fear arises when one's cooperation could be exploited by others' free-riding, while greed is the feeling associated with the temptation to free-ride on others' cooperation. According to social psychology and evolutionary theories, the author expected women to defect more when fear incentives were present, and men to defect more with greed incentives. Although he could corroborate only the second hypothesis, his intuition about women being more sensitive to the fear of being exploited is confirmed by other studies. For example, women level of cooperation is predicted by their trust in their interaction partner, while men's trust and cooperation are not linked [@Irwin2015]. Moreover, women's lower cooperation is explained by the fact that they expected their partner to contribute less to the common good [@Dorrough2019].

Finally, some of the analyzed papers presented interesting secondary results. First, against their authors' expectations, some intercultural studies showed that the gender difference in cooperation is not explained by the level of gender equality of the country [@Dorrough2019; @Spadaro2023]. Secondly, in an experiment not involving a social dilemma, women were significantly more altruistic than men [@Dorrough2019]. According to the authors, this underlines the importance of distinguishing between prosociality and strategic interactions that characterize social dilemmas. Therefore, in a situation where classical attributes of social dilemmas are not explicit, gender difference in cooperation could not be subjected to the same moderator effects described above. Lastly, according to @Sell2009, it is not possible to think about gender without relying it to social status, with men representing the high-status group. Therefore, gender differences in cooperation could also be caused by social status: to have a better understanding of this hypothesis, it would be interesting to investigate the difference in cooperation between two other social groups with a different status (e.g. high and low socioeconomic status).

In conclusion, "at times women are more cooperative than men, less cooperative than men, or equally as cooperative as men. These variations are explained, in large part, by structural differences and identities triggered by those differences" [@Sell2009]. The context in which men and women interact is therefore essential to predict if a gender will cooperate more than the other, and which one.

## Group identity

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group [e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020]. For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high identified individuals cooperated more than low identified ones when faced to an in-group participant who defected (and this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption, while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

However, despite the observed beneficial effects of inducing an intergroup competition, in a real-life situation this could be harmful. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by the competitive context, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

Finally, social identity does not only have a main effect on people's cooperation, but can also moderate other factors' impact. For example, monetary incentive are more effective at enhancing cooperation of low identified individual than of high identified ones [@VanVugt2009]. Moreover, a previous group failure can decrease cooperation between low identified member, but this effect is not observed in people who have a strong identification with their group [@Jackson2011; @Jackson2012].

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt2009]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful technique to enhance cooperation in an easy and cheap way.

## Values and personality traits

Personal values and personality traits can play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation is very consensual: in a meta-analysis, 81 out of the 82 papers examined reported than pro-social cooperated significantly more that pro-self [@Balliet2009].

However, this general effect is subjected to several moderators. For example, pro-self individuals cooperated more in a game where incentives to cooperation were present than in a game without such incentives, while pro-social did not displayed this difference and showed high levels of cooperation in both contexts [@Emonds2011]. A similar effect is also described by @Balliet2009, who found a stronger effect of SVO in unpaid games than in paid ones, because, when monetary rewards were present, pro-self increased their cooperation. Therefore, pro-self individuals also cooperate, but only when cooperation serves their personal interest. In fact, in a study about commuting habits, the extent to which the cooperative behaviour (i.e. choosing to commute by public transportation or carpooling) was perceived as efficient impacted pro-self individual's decision more than pro-social's [@VanLange1998].

Moreover, SVO has a stronger effect if there is an uncertainty about the number of individuals having access to a common resource: when the number of other participants was not known, pro-social, but not pro-self, reduced their harvesting comparing to a group-size certainty condition [@DeKwaadsteniet2008]. The same authors demonstrated that this result is explained by the fact that pro-social individuals want to maximize the joint outcome, while pro-self only think at their own personal interest. Finally, SVO effect is partly explained by trust, i.e. the extent to which people think that their interaction partners will also cooperate. In fact, in @VanLange1998, pro-social individuals with a low level of trust did not make different decisions than pro-self individuals, while the most cooperative were pro-social with an high level of trust.

Other studies also focused on how much people are self- or social-oriented by using other constructs. For example, @Liu2009a demonstrated that priming a common fate to induce a context of interdependence significantly increased the cooperation level of participants with an independent self, but had no effect on participants with an interdependent self (who already displayed high levels of cooperation). Moreover, in a social dilemma experiment, competitive, individualistic people were motivated just by personal gain, while people who saw themselves as a part of a group were more interested on maximizing the income of each member, even at the cost to have a smaller personal payoff [@Probst1999]. Also, high allocentric (i.e collectivist) individuals cooperated more than low allocentric ones when the perceived social norms are cooperative: this means that they are more sensitive to social cues and therefore they adapt their behaviour to social norms [@Chen2007].

Finally, there are other personal values and traits that can affect cooperation. For example, pro-environmental values are negatively correlated with harvesting from a common resource, but only when the resource is scarce; therefore these values are more likely to motivate difficult, rather than easy, ecological behaviours [@Sussman2016]. Levels of harvesting from a common resource are also lower between left-wingers than right-wingers [@Fosgaard2019]. Moreover, humility and honesty are significant predictor of cooperation [@Kieslich2014], and this is due to the fact that people with such characteristics manage to resist to the temptation to exploit others' contributions [@Hilbig2018]. Lastly, self-control also plays a role, in the sense that more self-control leads to a greater cooperation [@Kocher2017].

In conclusion, values and personality traits have a large impact on human cooperation, and it is thus very important to take them into account when investigating this phenomenon. Interestingly, some social psychological techniques (e.g. elaboration) allow to modify an individual's values [@Sussman2016]. Such techniques could therefore be used to improve cooperation.

## Social status???

retrieve work on gender that says that gender effects are probably in part impacted by status?

# Group factors

## Reward and punishment

Possible introduction: Incentives are divided in reward and punishments. Punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2016]. The positive effect of punishment in promoting cooperation is well confirmed ...

On the other side, results about rewarding are not so consensual ...

@Chen2022: reward, but not punishment, increased cooperation in the following round for both high and low benefit members

@Nockur2021: punishment (three different types) increase cooperation + the presence of a punishment system increased the perception of fairness + satisfaction was higher in peer and democratic punishment than in centralized and no punishment condition

@Kingsley2016: the possibility of punish others increased the cooperation in homogeneous groups (= where all members started with the same endowment), but not in heterogeneous groups

@Milinski2012: punishment is worthy on the long run (straight after, the ones who punished are the ones with the lowest income) + reward is less efficient than punishment (punish: at the beginning, to establish cooperation, reward is a cost all along because once cooperation is established you always have to continue rewarding it) + punishment can however entail a second-order social dilemma (defection = do not punish but having advantages for the fact that everybody cooperates thanks to punishers)

@Xiao2016: higher cooperation if only the cooperator can punish the defector (vs everybody can punish), and in particular for deterministic (vs stochastic) game\
when both can punish, anti-social punishment is present --\> effect = a person who cooperated and received a punishment in t is less likely to cooperate in t+1 (proportional)\
in stochastic games, the occurrence of a negative event is not a predictor for punishing (= based on other's defection, independently of the outcome)\
punishment of non-investor is smaller in stochastic (vs deterministic) games

@Balliet2011a: PUN: positive effect on cooperation in social dilemmas\
stronger if the punishment was costly (vs free)\
stronger if decentralized (vs centralized)\
stronger if SD with partners (vs strangers)\
stronger if iterated (vs one-shot)\
REWARD: positive effect on cooperation in social dilemmas\
stronger if costly vs free\
no effect of de-centralized\
stronger if iterated (vs one-shot)\
stronger in the resource dilemma (vs public good or prisoner dilemma)\
OVERALL: no difference on the strength of these effects

@Jacquet2011: the expectation of honour and shame enhance cooperation

@Rockenbach2019: successful groups punish, but with caution and not in early stages (see Excel for more results)

@VanDijk2016: results are both in the introduction and result case

exclusion as a form of punishment

## Communication

## Group size

## Social Norms

# Discussion

observation: majority of papers included are about SD, and very few studies about commons are quantitative --\> future research: quantitative approach to investigate cooperation in common initiatives

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
