---
title             : "Improving the understanding of cooperation in commons through social dilemmas' literature: A scoping review"
shorttitle        : "Improving the understanding of cooperation in commons"

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    #corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
     # - "Conceptualization"
     # - "Writing - Original Draft Preparation"
     # - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Centre for Development and Environment, University of Bern"
    
#authornote: |

abstract: |
  Commons represent an alternative social organization outside the classical capitalist model, therefore they could be the needed change toward more sustainable societies and economy. However, in order for a common to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in commons context. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since the two contexts share some structural characteristics, the results of social dilemma experiments could be used to gain insights on cooperation in commons.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, to gain a better understanding of what could shape cooperation in commons. A key-word search and a subsequent selection lead to the inclusion of 89 papers, who are used to identify 9 factors that can influence cooperation both at the individual and group level: *gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity*. These factors are discussed within the existing literature on commons and cooperation. *More specifically, the influences that these elements can have on each-other are highlighted, showing that the totality of factors should be taken into account when investigating cooperation in commons.* Finally, the practical implication of this work for commons' organization, as well as the review's limitations, are pointed out.
  
#keywords          : "cooperation, social dilemmas, commons"
#wordcount         : ""

bibliography      : CommonPaths.json

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
always_allow_html : true
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

**TO DO LIST:**

-   **try to shorten as much as possible the result section by putting detailed results on the table**

-   **put the references in the table (ex: gender, 11 papers --\> put the 11 references in the case)**

-   **look for journals and their requirements to know how much shorten and how to frame the introduction + discussion (ask Vincent which journal he chose?)**

# Individual factors

## Gender

The effect of gender on cooperation is controversial, and studies that investigated it provided contradictory results. Sometimes men cooperated more [@Dorrough2019], sometimes women did [@Peshkovskaya2017], but, in general, meta-analysis found a non significant effect size [e.g. @Balliet2011; @Spadaro2023]. This indicates that gender differences in cooperation are very flexible and can vary depending on the context [@Spadaro2023]. Therefore, when a significant effect is observed, it means that gender has somewhat been made salient by the situation [@Sell2009].

Interaction partners can moderate this effect. In fact, in a social dilemma experiment, women were more cooperative toward strangers than men, while men cooperated more in a group of friends. Moreover, after a socialization phase where strangers could interact, men cooperated more than women [@Peshkovskaya2017]. These results seem to indicate that men accord more importance at the relationship with other members than women. In fact, the creation of a team identity enhances men's, but not women's, cooperation [@Gomez-Ruiz2020]. Moreover, in a social dilemma with many iterations, men, but not women, tended to increase their cooperation over time [@Balliet2011].

The partners' gender also have an impact on women's and men's cooperation. In general, groups composed by men only are more cooperative than only-women groups. This is explained by the fact that women, compared to men, are generally less concerned by the maximization of their personal payoff and more about relative payoff. Thus, the defection strategy allows them to avoid earning less than the other players [@Colman2018]. Moreover, men display higher levels of gratitude and trust in same-sex groups than women [@Peshkovskaya2019]. On the other side, women cooperate more in mixed group. In fact, when both men and women are present, gender stereotypes are activated and people tend to adapt their behaviour to them. This leads women to behave in a more pro-social, altruistic way and men to be more self-oriented and competitive [@Balliet2011]. The same result is also found by @Peshkovskaya2019, who additionally demonstrated that women reported a stronger social connection with their partners in mixed than in same-sex groups. The overall group composition also have an impact on cooperation: in general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023]. However, @Barrero-Amortegui2021 showed that groups with a majority of women were more cooperative only after the signature of an agreement against excessive harvesting of a common resource. Therefore, even the effect of group composition seems to be subjected to moderator variables.

Another reason why gender differences in cooperation are often very small is that fear and greed motives are confused in a classical social dilemma situation [@Simpson2003]. In this contexts, a feeling of fear arises when one's cooperation could be exploited by others' free-riding, while greed is the feeling associated with the temptation to free-ride on others' cooperation. According to social psychology and evolutionary theories, the author expected women to defect more when fear incentives were present, and men to defect more with greed incentives. Although only the second hypothesis was corroborated, his intuition about women being more sensitive to the fear of being exploited is confirmed by other studies. For example, women level of cooperation is predicted by their trust in their interaction partner, while men's trust and cooperation are not linked [@Irwin2015]. Moreover, women's lower cooperation is explained by the fact that they expected their partner to contribute less to the common good [@Dorrough2019].

Finally, some of the analyzed papers presented interesting secondary results. First, against their authors' expectations, some intercultural studies showed that the gender difference in cooperation is not explained by the level of gender equality of the country [@Dorrough2019; @Spadaro2023]. Secondly, in an experiment not involving a social dilemma, women were significantly more altruistic than men [@Dorrough2019]. According to the authors, this underlines the importance of distinguishing between prosociality and strategic interactions that characterize social dilemmas. Therefore, in a situation where classical attributes of social dilemmas are not explicit, gender difference in cooperation could not be subjected to the same moderator effects described above. Lastly, according to @Sell2009, it is not possible to think about gender without relying it to social status, with men representing the high-status group. Therefore, gender differences in cooperation could also be caused by social status: to have a better understanding of this hypothesis, it would be interesting to investigate the difference in cooperation between two other social groups with a different status (e.g. high and low socioeconomic status).

In conclusion, "at times women are more cooperative than men, less cooperative than men, or equally as cooperative as men. These variations are explained, in large part, by structural differences and identities triggered by those differences" [@Sell2009, p. 322]. The context in which men and women interact is therefore essential to predict if a gender will cooperate more than the other, and which one.

## Social status

Few studies have investigated how social status impact cooperation, and their results do not allow to draw a clear conclusion about this effect. For instance, advantaged members (i.e. high-status class) tend to cooperate less than disadvantaged ones [@Aksoy2019]. In fact, when participants did not receive the same endowment at the beginning of the experiment, contributions of "rich" and "poor" members converged [@Kingsley2016]. In other words, proportionally to their resources, low-status participants cooperated more. However, in their review, @VanLange2013 identified some studies that observed the opposite effect, namely that wealthier people (or those who benefited more from a common resource) contributed more to the common good than the other.

Social status can also create inter-class conflicts. In fact, people cooperated more with the members of the same social class than with other classes' members [@Aksoy2019]. Moreover, a form of solidarity is observed between low-status members: in a study where participants had the opportunity to punish others for a non-cooperative behaviour, low-status members punished significantly less other low-status participants [@Chen2022]. Also, in another experiment, participants were randomly advantaged during the interaction: subjects who received a lot of advantages were discriminated by others, and especially by the ones who were rarely advantaged [@Camera2020].

Regarding the overall group composition, results are not consensual. On one hand, @Aksoy2019 found that groups with an inequality of status between members displayed lower levels of cooperation than homogeneous groups. On the other hand, economic and socio-cultural heterogeneity, although being negatively correlated with in-group trust, does not have a direct effect on a common-pool resource well-being [@VanKlingeren2021].

In conclusion, social status impacts cooperation, but it seems to have multiple effects. Moreover, similarly to what @Sell2009 affirmed about the effects of gender and social status, @Aksoy2019 demonstrated that the lower level of cooperation between members of different social classes is mainly due to actors' identification with their classes. This shows that social identity and social status are deeply connected, and that it is difficult to separate their effects on cooperation.

## Group identification

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group [e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020]. For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Group identification can also be easily implemented in a laboratory experiment, and it has still a strong positive effect on cooperation [@Wit1992]. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high-identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high-identified individuals cooperated more than low-identified ones when faced to an in-group participant who defected (and this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In fact, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption, while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition in itself.

However, despite the observed beneficial effects of inducing an intergroup competition, this could be harmful in real-life situations. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by the competitive context, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

Finally, social identity does not only have a main effect on people's cooperation, but can also moderate other factors' impact. For example, monetary incentive are more effective at enhancing cooperation of low-identified individual than of high-identified ones [@VanVugt2009]. Moreover, a previous group failure can decrease cooperation between low-identified member, but this effect is not observed in people who have a strong identification with their group [@Jackson2011; @Jackson2012].

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt2009]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful technique to enhance cooperation in an easy and cheap way.

## Values and personality traits

Personal values and personality traits can play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation is very consensual: in a meta-analysis, 81 out of the 82 papers examined reported that pro-social cooperated significantly more than pro-self [@Balliet2009].

However, this general effect is subjected to several moderators. For example, pro-self individuals cooperated more in a game where economic incentives to cooperation were present than in a game without such incentives, while pro-social did not display this difference and showed high levels of cooperation in both contexts [@Emonds2011]. A similar effect is also described by @Balliet2009, who found a stronger effect of SVO in unpaid games than in paid ones, because, when monetary rewards were present, pro-self increased their cooperation. Therefore, pro-self individuals also cooperate, but only when cooperation serves their personal interest. In fact, in a study about commuting habits, the extent to which the cooperative behaviour (i.e. choosing to commute by public transportation or carpooling) was perceived as efficient impacted pro-self individual's decision more than pro-social's [@VanLange1998].

Moreover, SVO has a stronger effect if there is an uncertainty about the number of individuals having access to a common resource: when the number of other participants was not known, pro-social, but not pro-self, reduced their harvesting comparing to a group-size certainty condition [@DeKwaadsteniet2008]. The same authors demonstrated that this result is explained by the fact that pro-social individuals want to maximize the joint outcome, while pro-self mainly think at their own personal interest. Also, SVO effect is partly explained by trust, i.e. the extent to which people think that their interaction partners will also cooperate. In fact, in @VanLange1998, pro-social individuals with a low level of trust did not make different decisions than pro-self individuals, while the most cooperative were pro-social with an high level of trust. Finally, the intuitive choice is to cooperate for pro-social individuals, while it is to defect for pro-self [@Lu2019]. However, pro-self can cooperate more if they have the possibility to deliberate about the situation, because they realize that the best strategy to maximize the joint (and thus personal) outcome is cooperation.

Other studies also focused on how much people are self- or social-oriented by using other constructs. For example, @Liu2009a demonstrated that priming a common fate to induce a context of interdependence significantly increased the cooperation level of participants with an independent self, but had no effect on participants with an interdependent self (who already displayed high levels of cooperation). Moreover, in a social dilemma experiment, competitive, individualistic people were motivated just by personal gain, while people who saw themselves as a part of a group were more interested on maximizing the income of each member, even at the cost to have a smaller personal payoff [@Probst1999]. Also, high allocentric (i.e collectivist) individuals cooperated more than low allocentric ones when the perceived social norms are cooperative: this means that they are more sensitive to social cues and therefore they adapt their behaviour to social norms [@Chen2007].

Finally, there are other personal values and traits that can affect cooperation. For example, pro-environmental values are negatively correlated with harvesting from a common resource, but only when the resource is scarce; therefore these values are more likely to motivate difficult, rather than easy, ecological behaviours [@Sussman2016]. Levels of harvesting from a common resource are also lower between left-wingers than right-wingers [@Fosgaard2019]. Moreover, humility and honesty are significant predictors of cooperation [@Kieslich2014], and this is due to the fact that people with such characteristics manage to resist to the temptation to exploit others' contributions [@Hilbig2018]. Self-control also plays a role, in the sense that more self-control leads to a greater cooperation [@Kocher2017]. Lastly, people with an higher justice sensitivity make bigger contributions to a public good [@Schlosser2018].

In conclusion, values and personality traits have a large impact on human cooperation, and it is thus very important to take them into account when investigating this phenomenon. Interestingly, some social psychological techniques (e.g. elaboration) allow to modify an individual's values [@Sussman2016]. Such techniques could therefore be used to improve cooperation.

## Trust and beliefs about others' contribution

Trust can be defined as the "expectations of others' benevolent motives in situations that involve a conflict between self and collective interests" [@Balliet2013, p.1]. Trust can be dispositional or a state; the former is an individual's general trait, and is usually measured with personality scales, the latter is linked to the context, and can be operazionalized by expectations about a partner's behaviour in a specific situation [@Balliet2013]. The two are positively correlated [@Lubke2021], and both have a positive general effect on cooperation [@VanLange2013; @Balliet2013]. In fact, high levels of trust are associated with more contributions to a public good [@Kocher2015], and a participant's contribution in a public good game can be predicted by her beliefs about others' contributions [@Smith2013].

Several studied confirmed this relationship. For instance, @Chaudhuri2002 demonstrated that people who cooperated more in a first public good game trusted more their partner in a following interaction than people whose main strategy was defecting. Moreover, people that were pessimistic about others' contribution cooperated significantly less than the optimistic [@Oyediran2018], and those who free-rode in a public good game, compared to cooperators, believed that others contributed less [@Kocher2015]. Also, people were more willing to purchase actions for a CO2 mitigation initiative when they believed that the majority of others had also bought them [@Sturm2019]. Finally, in a study about self-administered outlets in Swiss farms (i.e. outlets where nobody monitors if people pay or not their purchases), @Franzen2019 found that a farmer's general trust in other people is a significant predictor of the probability of having such an outlet. Concerning the moderators, trust matters more in situations involving a large, rather than small, conflict between personal and collective interests [@Balliet2013], and when people are uncertain about others' intentions [@VanLange2013].

People can also display different levels of trust in the "reliability, effectiveness and legitimacy of public institutions" [@Rompf2017, p. 2]. These authors showed that institutional trust is associated with the probability of recycling, because knowing that sanctions can be efficiently applied can motivate people to comply to norms. Moreover, the effects of perceived costs and benefits of the action are less predictive in people with high institutional trust.

Finally, trusting others (and the institutions) is essential to display higher level of cooperation. However, general beliefs about others' cooperation are quite low compared to the actual behaviours, and this trust gap can undermine cooperation, since people's actions are significantly predicted by their beliefs [@Lubke2021]. In order to increase trust and, consequently, cooperation, it is therefore essential to provide people with reliable information about others' behaviour.

# Organizational factors

## Incentives

Cooperation rates can be modified by introducing incentives, who can be either punishments or rewards[^1]. In the literature, punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2015], and its positive effect on cooperation is very consensual. In general, when there is the possibility for participants to punish others members based on their contributions, cooperation is higher than when this is not possible [e.g. @Pfattheicher2018; @Balliet2011a]. Moreover, the presence of a punishment system increases the perception of fairness [@Nockur2021].

[^1]: Since the interest of the present review is to investigate how the internal organization of a group can affect cooperation, only papers that studied situations where the participants could give rewards or punishments to each-others are included. Incentives given by third party have also an effect on cooperation, but they are not included in this review.

However, this main results is subjected to several moderators. First, the punishment has a bigger impact on cooperation if its implementation is costly for the punishers than if it is free [@Balliet2011a], but at the same time the cost should not be too high [@Egas2008]. Moreover, punishments are more effective if the interaction is among friends than among strangers, and if people know that they will have to interact repeatedly with each other [@Balliet2011a]. In fact, in short interactions, if punishment is costly, individuals who punish others are the ones with the lowest personal income [@Milinski2012], and this could discourage the application of the fine. Also, an experiment showed that a punishment system increased cooperation in groups where all members had the same amount of resources, but not in groups where income was heterogeneous [@Kingsley2016]. In fact, @Chen2022 found that, in groups where the initial endowment was heterogeneous, participants that received less didn't punish the others "low-status" members. This solidarity could therefore encourage free-riding. Moreover, the effect of punishment is stronger when only cooperators have the power to punish defectors, than when everybody can punish [@Xiao2016]. Finally, a plain situation is also needed to have a stronger effect of punishment: the link between behaviour and outcome must be clear, without random factors affecting the final outcome [@Xiao2016], and the punishment must be highly effective [@Egas2008].

However, punishment must be applied with carefulness: @Rockenbach2019 compared the strategies of different groups, and found that the most successful ones (whose members had the highest payoff at the end of the experiment) punished with caution and not in early stages of the game. In these groups, a norm of when and how punishment should be applied seems to exist. Moreover, punishment has also two main negative sides. First, it could entails a second-order social dilemma [@Milinski2012], since cooperative participants can free-ride on the effort of other cooperators who punish non-cooperators. Secondly, anti-social punishment can arise. This term refer to the punishment of cooperators by non-cooperators, and it is present when all participants can punish independently of their contributions [@Xiao2016]. In this context, defectors can see cooperators as rule breakers, and anti-social punishment is a sort of retaliation [@VanDijk2015].

A possible solution could be to implement a system of democratic punishment, where a fine is given only if a majority of participants agree. In fact, compared to a peer punishment system where participants decide individually if and who punish, in a democratic system anti-social punishment is less present, general punishment level is lower, and final payoff and cooperation rates are higher [@Pfattheicher2018].

Rewarding systems, on the other side, are less investigated. Globally, they have a positive effect on cooperation [@Balliet2011a]. However, some studies found that they were more efficient than punishment [e.g. @Chen2022], while others came to the opposite conclusion [e.g. @Milinski2012] or found no difference between the two systems [e.g. @Balliet2011a]. In a meta-analysis, @Balliet2011a found that rewards had a stronger effect on cooperation if their implementation was costly rather than free, and in iterated, rather than single, interactions. Moreover, rewarding was more effective in situations where participants had to decide how much take from a common resource, compared to when they had to decide how much contribute to a common good.

In general, incentives seem to be a powerful tool to enhance cooperation, but they have a dark side. First, they reduce participants' intrinsic motivation [@VanDijk2015], by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. In fact, @Schlosser2018 demonstrated that people's level of justice sensitivity could predict their contributions to a public good game, but this link disappeared when a punishment system was introduced. Secondly, when incentives are present people think that others cooperate just because of them, and this undermines their trust in other members, and especially the trust in the fact that others are internally motivated to cooperate. This reduced trust in others drops cooperation levels if incentives are removed [@Irwin2014; @Mulder2006].

Finally, incentives are often studied in economic formats, but they can also be social. For example, the expectation of honour or shame (after cooperating or defecting, respectively) is enough to enhance cooperation [@Jacquet2011]. Moreover, expressing disapproval toward defectors can boost cooperation by inducing more compliance to the cooperative social norm [@Nelissen2013]. Even an ostracism system where participants can vote to exclude other members from the game is very effective in increasing cooperation [@Maier-Rigaud2010]. In general, social sanctions can be even more efficient than economic ones, because cooperation do not drop after their removal [@Nelissen2013].

To conclude, economic incentives can have a positive effect on cooperation, but their efficient implementation must be insured all along the interaction, otherwise cooperation rates would drop. On the other hand, social incentives seem to have the same positive effect, and their removal do not cause a reduction of cooperation. However, there is a lack of studies focusing on this kind of sanctions.

## Communication

Communication has a global positive effect on cooperation [e.g. @Balliet2010; @Bicchieri2002], but this effect is subjected to several moderators. First, talking is more powerful than sending written messages, because it allows non-verbal communication and facilitates the social norms formation process. Moreover, communication's effect is stronger in large groups: when many people are present cooperation rates are lower, so communication between members can buffer this negative effect [@Balliet2010]. Finally, if communicating with others requires a personal cost, the effect is weaker than if communication is free [@Janssen2014].

The content of the messages is also essential. @Koessler2021 identified four steps necessary for an effective communication in social dilemmas: problem awareness, identification of strategies, agreement and ratification. In another study, groups who used communication to foster a common social identity, promote cooperative social norms and ensure that all the players publicly committed displayed higher rates of cooperation than other groups [@Adams2022]. However, casual chit-chat also plays a very important role in increasing cooperation [@Koessler2021]. Also, the increase of cooperation through communication can be explained by the normative content that people are shaping while exchanging messages [@Janssen2014], and by the fact that it allows participant to reach a consensus about the best strategy [@Hopthrow2005]. Moreover, even if communication also helps to create a social identity, the effect of social norms is a stronger mediator of the link between communication and cooperation [@Bicchieri2002].

Another factor that influences the effect of communication on cooperation is whether individuals exchange messages before, during or after the cooperative phase. Concerning the post-game communication, @Kumakawa2013 found that receiving a feedback about the cooperation choice made previously can increase cooperation in the following interactions, especially if the feedback is negative. However, just the fact of knowing that a discussion will follow is not enough to enhance cooperation, and can even have a counterproductive effect, because forcing people to think about a future discussion make them reflect more on the situation and thus act more rationally [@Torsvik2011]. On the other hand, communication before or during the interaction have both a positive impact on cooperation, and the difference between their effect is not significant [@Balliet2010].

Finally, several studies investigated the effect of gossip as a form of communication that can increase cooperation. Gossip is defined as a spread of reputational information about another person [@Feinberg2014], and it increases cooperation in two ways. First, if people know that gossip is possible, they tend to cooperate more to avoid the risk of being the target of such gossip and therefore of being excluded from following interactions. Secondly, having information about the past behaviour of potential partners can help to select the most cooperative individuals in an interaction situation [@Feinberg2014]. Moreover, the same authors found that pro-social individuals gossiped more than pro-self. This shows that gossip can have the pro-social function of protecting others from a potential exploitation [@Feinberg2012]. However, gossip has also some self-interest effects: for instance, observing another subject being exploited and having the opportunity to gossip about the exploiter significantly reduced negative affects in @Feinberg2012 's participants. Moreover, providing correct information while gossiping helps to build a positive reputation [@Giardini2021].

In conclusion, when people have the possibility to communicate before, during or after a cooperative interaction, the cooperation level increases. Even if the strength of this effect can vary depending on the situation, communication never seems to have a negative impact on cooperation.

## Social Norms

Social norms can provide a guide on how to behave, especially when there is uncertainty about others' behaviour [@VonBorgstede2018]. @Thogersen2008 studied the effect of social norms on daily social dilemma situations such as environmental behaviours, and found that both injunctive (what others approve) and descriptive (what others do) norms are positively related to behaviours. Moreover, cooperation can be maintained via social norms, because people who do not respect them are susceptible of social sanctions. In fact, if social approval is possible, people cooperate more [@Rege2004].

There are several explanations of social norms' effect on cooperation. First, according to the appropriateness framework, when taking a decision people ask themselves: "What does a person like me do in a situation like this?" [@Arora2012]. Therefore they need to define the situation as a cooperative task and to strongly identify with the group [@VanLange2013]. This leads to a greater accessibility of group's norms and consequently (if the norms are cooperative) to more cooperation. Testing this assumption, @Arora2012 showed that a greater identification with social group lead to higher levels of cooperation. The second explanation concerns conformity: in general people adapt their cooperation to others' behaviour, but this is usually explained by the reciprocity principle (cooperation enhances the probability that others will do the same, and therefore that the final payoff will be higher) [@Bardsley2005a]. However, these authors demonstrated that people conform to others' behaviour even if this has no impact on their personal payoff. Therefore, people have a general tendency to imitate others to follow social norms.

Finally, social norms can be modified by participants. For instance, @Weber2008 demonstrated that in several social dilemma experiment there are consistent cooperators, i.e. people who always cooperate. The presence of these people can increase other members' cooperation, and this effect is mediated by the perception of the social norms. Moreover, social norms can also be shaped via communication [@Bicchieri2002].

In conclusion, social norms have a significant effect on people's cooperation in social dilemma situations. Since they are quite easily modified [e.g. @Weber2008], shaping cooperative norms can be a powerful instrument to improve cooperation.

## Anonymity

There are not many studies on the influence of anonymity on cooperation, but, according to @Wang2017, most of the factors that can increase cooperation rely on a reduction of anonymity. For instance, it is not possible to punish defectors without knowing who they are. Therefore, in general it seems that non-anonymous situations can significantly increase cooperation and reduce defection [@Wang2017]. This can be explained by the fact that, when their actions are visible, people worry about others' approval [@Rege2004] and that, in non-anonymous situations, contributions can be a self-presentation strategy, because being perceived as pro-social can increase one's social status [@VanVugt2010a]. There are therefore more social incentives to cooperate in non-anonymous, rather than anonymous, situations.

Moreover, this effect is displayed especially in situations when individuals face a cost when cooperating and know that they will not get any personal benefit. For instance, @Butz2021 demonstrated that, in a public good game where a part of contributions was donated to a charity association, disclosing identities increased cooperation, but only if the donation was deducted from participants' payoff. In the same direction, @VanVugt2010a found that, when actions were public, contributions were higher if the public good was already attained or unattainable, and therefore if participants did not have a personal interest in cooperating.

Reducing anonymity in cooperative interactions can be very effective to encourage people to contribute more. A possible way to do this is to reduce group's size. In fact, according to @Romano2016, larger groups lead to de-individuation, because it is more difficult to understand who is defecting. Therefore, it is easier for members to feel anonymous and to perceive the social punishments associated with defection as less probable. Moreover, communication is easier in smaller groups [@VonBorgstede2018]. This assumption is confirmed experimentally by @Jiang2021, who showed than in smaller groups people contributed more often to a public good.

Another effective method to reduce anonymity is to make people publicly promise to cooperate before the cooperative interaction. For instance, in a laboratory experiment modeling a common pool resource, people who publicly committed exploited significantly less the resource than people who committed in private, even if after committing they were not restricted in their harvesting possibilities [@Mosler1993]. This results was also confirmed by @Przepiorka2020, who showed that non-binding promises increased cooperation. Moreover, even a compulsory commitment has a beneficial effect on cooperation, even if the most efficient groups are those where all participants signed voluntary the pledge [@Koessler2021a].

In conclusion, reducing anonymity during interactions is essential to improve cooperation, because this can facilitate the implementation of others cooperation-enhancing techniques [@Wang2017]. Moreover, there are several low-cost method to make situations non-anonymous.

# 

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
