---
title             : "Improving the understanding of cooperation in commons through social dilemmas' literature: A scoping review"
shorttitle        : "Improving the understanding of cooperation in commons"

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    #corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
     # - "Conceptualization"
     # - "Writing - Original Draft Preparation"
     # - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Centre for Development and Environment, University of Bern"
    
#authornote: |

abstract: |
  Commons represent an alternative social organization outside the classical capitalist model, therefore they could be the needed change toward more sustainable societies and economy. However, in order for a common to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in commons context. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since the two contexts share some structural characteristics, the results of social dilemma experiments could be used to gain insights on cooperation in commons.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, to gain a better understanding of what could shape cooperation in commons. A key-word search and a subsequent selection lead to the inclusion of 89 papers, who are used to identify 9 factors that can influence cooperation both at the individual and group level: *gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity*. These factors are discussed within the existing literature on commons and cooperation. *More specifically, the influences that these elements can have on each-other are highlighted, showing that the totality of factors should be taken into account when investigating cooperation in commons.* Finally, the practical implication of this work for commons' organization, as well as the review's limitations, are pointed out.
  
#keywords          : "cooperation, social dilemmas, commons"
#wordcount         : ""

bibliography      : CommonPaths.json

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
always_allow_html : true
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

**TO DO LIST:**

-   **try to shorten as much as possible the result section by putting detailed results on the table**

-   **put the references in the table (ex: gender, 11 papers --\> put the 11 references in the case)**

-   **look for journals and their requirements to know how much shorten and how to frame the introduction + discussion (ask Vincent which journal he chose?)**

# Individual factors

Five individuals factors that can impact cooperation emerged from the litterature review: gender, social status, group identification, values and personality traits, and trust. In the following sections, the principal effects of these factors on cooperation are described. Moreover, an overview of the moderators of each effect is provided. More details about moderators are included in the Annexe A.

## Gender

The effect of gender on cooperation is controversial, and studies that investigated it provided contradictory results. Sometimes men cooperated more [@Dorrough2019], sometimes women did [@Peshkovskaya2017], but two meta-analysis found a non significant effect size [@Balliet2011; @Spadaro2023]. This indicates that gender differences in cooperation are very flexible and depend on the context [@Spadaro2023]. More specifically, @Sell2009 explained this contradictory results "by structural differences [of the experimental settings] and identities triggered by those differences" (p. 322).

The effect of gender on cooperation is therefore subjected to several moderators, since the situation in which women and men interact is essential to determine which gender will cooperate more. For instance, men are more cooperative when they know the other participants [@Peshkovskaya2017], when they feel part of a team [@Gomez-Ruiz2020] or when they interact repeatedly with the same group [@Balliet2011]. The partners' gender is also an important moderator. Men-only groups are more cooperative than women-only [@Colman2018; @Peshkovskaya2019], while women cooperate more in mixed groups [@Balliet2011; @Peshkovskaya2019]. In general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023], even if this effect of group composition can be moderated by other factos [e.g. @Barrero-Amortegui2021]. Moreover, men are more sensitive to greed incentives than women , and thus they cooperate less in situations with strong temptations to free-ride are present [@Simpson2003]. On the other hand, women's principal motive in cooperative interaction is avoiding to be exploited, and that is why the cooperate less if they don't trust other participants or if they believe than others provide small contributions [@Irwin2015; @Dorrough2019].

Lastly, gender differences are also linked to social status, with women representing the low-status group. Therefore, gender differences in cooperation could also be partly explained by social status: to have a better understanding of this hypothesis, it is important to understand how social status impact cooperation.

## Social status

The results of studies investigating the impact of social status on cooperation do not allow to draw a clear conclusion about this effect. For instance, some findings suggest that low-status participants cooperate more than high-status ones by contributing a higher proportion of their endowments [@Kingsley2016]. However, @VanLange2013 identified some studies where wealthier people cooperated more.

Social status can also create inter-class conflicts, because people tend to cooperate more with the members of the same social class than with other classes' members [@Aksoy2019]. This is especially true for low-status groups, where solidarity can arise either in the form of discrimination toward high-status members [@Camera2020], or of less punishments for other low-status [@Chen2022].

Regarding the overall group composition, results are not consensual. On one hand, @Aksoy2019 found that groups with an inequality of status between members displayed lower levels of cooperation than homogeneous groups. On the other hand, economic and socio-cultural heterogeneity, although being negatively correlated with in-group trust, does not have a direct effect on a common-pool resource well-being [@VanKlingeren2021].

Lastly, similarly to what @Sell2009 affirmed about the effects of gender and social status, @Aksoy2019 demonstrated that the lower level of cooperation between members of different social classes is mainly due to actors' identification with their classes. This shows that social identity and social status are deeply connected, and that it is difficult to separate their effects on cooperation.

## Group identification

Social identity have a widely demonstrated positive effect on cooperation: the more individuals identify with a group, the more they cooperate with other members of the group [e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020]. In fact, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goals [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. The strength of this effect is further confirmed by the fact that it is the same for individualist and collectivist individuals [@Chen2007], and that it works even when groups are created in a laboratory experiment [@Wit1992].





Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Group identification can also be easily implemented in a laboratory experiment, and it has still a strong positive effect on cooperation [@Wit1992]. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high-identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high-identified individuals cooperated more than low-identified ones when faced to an in-group participant who defected (and this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In fact, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption, while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition in itself.

However, despite the observed beneficial effects of inducing an intergroup competition, this could be harmful in real-life situations. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by the competitive context, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

Finally, social identity does not only have a main effect on people's cooperation, but can also moderate other factors' impact. For example, monetary incentive are more effective at enhancing cooperation of low-identified individual than of high-identified ones [@VanVugt2009]. Moreover, a previous group failure can decrease cooperation between low-identified member, but this effect is not observed in people who have a strong identification with their group [@Jackson2011; @Jackson2012].

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt2009]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful technique to enhance cooperation in an easy and cheap way.

## Values and personality traits

Personal values and personality traits play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation has been widely demonstrated: in a meta-analysis, 81 out of the 82 papers examined reported that pro-social cooperated significantly more than pro-self [@Balliet2009]. 

There are several moderators to this main effect. For instance, when economic incentives to cooperation are present, pro-self individuals attain a level of cooperation similar to pro-social's one [@Emonds2011; @Balliet2009]. Similarly, in a study about commuting habits, the extent to which the cooperative behaviour (i.e. choosing to commute by public transportation or carpooling) was perceived as efficient impacted pro-self individual's decision more than pro-social's [@VanLange1998]. Another moderator is the uncertainity about how many individuals compose the group: if the uncerainty is high pro-social, but not pro-self, increse their level of cooperation [@DeKwaadsteniet2008]. Finally, pro-self increase their cooperation if they have the possibility to deliberate about the situation [@Lu2019].

Other studies also focused on how much people are self- or social-orientation impact cooperation by using other constructs. More cooperative individuals have an interdependent definition of the self [@Liu2009a] and see themselves as part of the group [@Probst1999]. Moreover, allocentric (i.e. collectivist) people are more sensitive to cooperative social norms [@Chen2007]. Finally, other personal values and traits that affect cooperation include pro-environmental values [@Sussman2016], political orientation [@Fosgaard2019], humility and honesty [@Kieslich2014; Hilbig2018], self-control [@Kocher], and justice sensitivity [@Schlosser2018].

## Trust and beliefs about others' contribution

Trust can be defined as the "expectations of others' benevolent motives in situations that involve a conflict between self and collective interests" [@Balliet2013, p.1]. There are two category of trust: first, dispositional trust is an individual general trait that can be assessed with personality scale. Secondly, state trust is linked to the context and represent the expectations about a pertner's behaviour in a specific situation [@Balliet2013]. The two are positively correlated [@Lubke2021], and both have a positive general effect on cooperation [@VanLange2013; @Balliet2013].

Several studied confirmed this relationship: high levels of trust are associated with more contributions to a public good [@Kocher2015], individuals' cooperation can be predicted by their beliefs about others' contributions [@Smith2013], people who are pessimistic about others' contribution cooperate significantly less than optimistic [@Oyediran2018], and individuals who cooperate more display higher levels of trust in following interactions [@Chaudhuri2002]. The effect of trust on cooperation has also been demostrated in real-life studies [e.g. @Sturm2019; @Franzen2019]. Finally, trust matters more in situations involving a large, rather than small, conflict between personal and collective interests [@Balliet2013], and when people are uncertain about others' intentions [@VanLange2013].

People can also display different levels of trust in the "reliability, effectiveness and legitimacy of public institutions" [@Rompf2017, p. 2]. Institutional trust is associated with the probability of recycling, because knowing that sanctions can be efficiently applied can motivate people to comply to norms. Moreover, the effects of perceived costs and benefits of the action are less predictive in people with high institutional trust.

# Organizational factors

## Incentives

Cooperation rates can be modified by introducing incentives, who can be either punishments or rewards[^1]. In the literature, punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2015], and its positive effect on cooperation is very consensual. In general, when there is the possibility for participants to punish others members based on their contributions, cooperation is higher than when this is not possible [e.g. @Pfattheicher2018; @Balliet2011a]. Moreover, the presence of a punishment system increases the perception of fairness [@Nockur2021].

[^1]: Since the interest of the present review is to investigate how the internal organization of a group can affect cooperation, only papers that studied situations where the participants could give rewards or punishments to each-others are included. Incentives given by third party have also an effect on cooperation, but they are not included in this review.

However, this main results is subjected to several moderators. First, the punishment has a bigger impact on cooperation if its implementation is costly for the punishers than if it is free [@Balliet2011a], but at the same time the cost should not be too high [@Egas2008]. Moreover, punishments are more effective if the interaction is among friends than among strangers, and if people know that they will have to interact repeatedly with each other [@Balliet2011a]. In fact, in short interactions, if punishment is costly, individuals who punish others are the ones with the lowest personal income [@Milinski2012], and this could discourage the application of the fine. Also, an experiment showed that a punishment system increased cooperation in groups where all members had the same amount of resources, but not in groups where income was heterogeneous [@Kingsley2016]. In fact, @Chen2022 found that, in groups where the initial endowment was heterogeneous, participants that received less didn't punish the others "low-status" members. This solidarity could therefore encourage free-riding. Moreover, the effect of punishment is stronger when only cooperators have the power to punish defectors, than when everybody can punish [@Xiao2016]. Finally, a plain situation is also needed to have a stronger effect of punishment: the link between behaviour and outcome must be clear, without random factors affecting the final outcome [@Xiao2016], and the punishment must be highly effective [@Egas2008].

However, punishment must be applied with carefulness: @Rockenbach2019 compared the strategies of different groups, and found that the most successful ones (whose members had the highest payoff at the end of the experiment) punished with caution and not in early stages of the game. In these groups, a norm of when and how punishment should be applied seems to exist. Moreover, punishment has also two main negative sides. First, it could entails a second-order social dilemma [@Milinski2012], since cooperative participants can free-ride on the effort of other cooperators who punish non-cooperators. Secondly, anti-social punishment can arise. This term refer to the punishment of cooperators by non-cooperators, and it is present when all participants can punish independently of their contributions [@Xiao2016]. In this context, defectors can see cooperators as rule breakers, and anti-social punishment is a sort of retaliation [@VanDijk2015].

A possible solution could be to implement a system of democratic punishment, where a fine is given only if a majority of participants agree. In fact, compared to a peer punishment system where participants decide individually if and who punish, in a democratic system anti-social punishment is less present, general punishment level is lower, and final payoff and cooperation rates are higher [@Pfattheicher2018].

Rewarding systems, on the other side, are less investigated. Globally, they have a positive effect on cooperation [@Balliet2011a]. However, some studies found that they were more efficient than punishment [e.g. @Chen2022], while others came to the opposite conclusion [e.g. @Milinski2012] or found no difference between the two systems [e.g. @Balliet2011a]. In a meta-analysis, @Balliet2011a found that rewards had a stronger effect on cooperation if their implementation was costly rather than free, and in iterated, rather than single, interactions. Moreover, rewarding was more effective in situations where participants had to decide how much take from a common resource, compared to when they had to decide how much contribute to a common good.

In general, incentives seem to be a powerful tool to enhance cooperation, but they have a dark side. First, they reduce participants' intrinsic motivation [@VanDijk2015], by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. In fact, @Schlosser2018 demonstrated that people's level of justice sensitivity could predict their contributions to a public good game, but this link disappeared when a punishment system was introduced. Secondly, when incentives are present people think that others cooperate just because of them, and this undermines their trust in other members, and especially the trust in the fact that others are internally motivated to cooperate. This reduced trust in others drops cooperation levels if incentives are removed [@Irwin2014; @Mulder2006].

Finally, incentives are often studied in economic formats, but they can also be social. For example, the expectation of honour or shame (after cooperating or defecting, respectively) is enough to enhance cooperation [@Jacquet2011]. Moreover, expressing disapproval toward defectors can boost cooperation by inducing more compliance to the cooperative social norm [@Nelissen2013]. Even an ostracism system where participants can vote to exclude other members from the game is very effective in increasing cooperation [@Maier-Rigaud2010]. In general, social sanctions can be even more efficient than economic ones, because cooperation do not drop after their removal [@Nelissen2013].

To conclude, economic incentives can have a positive effect on cooperation, but their efficient implementation must be insured all along the interaction, otherwise cooperation rates would drop. On the other hand, social incentives seem to have the same positive effect, and their removal do not cause a reduction of cooperation. However, there is a lack of studies focusing on this kind of sanctions.

## Communication

Communication has a global positive effect on cooperation [e.g. @Balliet2010; @Bicchieri2002], but this effect is subjected to several moderators. First, talking is more powerful than sending written messages, because it allows non-verbal communication and facilitates the social norms formation process. Moreover, communication's effect is stronger in large groups: when many people are present cooperation rates are lower, so communication between members can buffer this negative effect [@Balliet2010]. Finally, if communicating with others requires a personal cost, the effect is weaker than if communication is free [@Janssen2014].

The content of the messages is also essential. @Koessler2021 identified four steps necessary for an effective communication in social dilemmas: problem awareness, identification of strategies, agreement and ratification. In another study, groups who used communication to foster a common social identity, promote cooperative social norms and ensure that all the players publicly committed displayed higher rates of cooperation than other groups [@Adams2022]. However, casual chit-chat also plays a very important role in increasing cooperation [@Koessler2021]. Also, the increase of cooperation through communication can be explained by the normative content that people are shaping while exchanging messages [@Janssen2014], and by the fact that it allows participant to reach a consensus about the best strategy [@Hopthrow2005]. Moreover, even if communication also helps to create a social identity, the effect of social norms is a stronger mediator of the link between communication and cooperation [@Bicchieri2002].

Another factor that influences the effect of communication on cooperation is whether individuals exchange messages before, during or after the cooperative phase. Concerning the post-game communication, @Kumakawa2013 found that receiving a feedback about the cooperation choice made previously can increase cooperation in the following interactions, especially if the feedback is negative. However, just the fact of knowing that a discussion will follow is not enough to enhance cooperation, and can even have a counterproductive effect, because forcing people to think about a future discussion make them reflect more on the situation and thus act more rationally [@Torsvik2011]. On the other hand, communication before or during the interaction have both a positive impact on cooperation, and the difference between their effect is not significant [@Balliet2010].

Finally, several studies investigated the effect of gossip as a form of communication that can increase cooperation. Gossip is defined as a spread of reputational information about another person [@Feinberg2014], and it increases cooperation in two ways. First, if people know that gossip is possible, they tend to cooperate more to avoid the risk of being the target of such gossip and therefore of being excluded from following interactions. Secondly, having information about the past behaviour of potential partners can help to select the most cooperative individuals in an interaction situation [@Feinberg2014]. Moreover, the same authors found that pro-social individuals gossiped more than pro-self. This shows that gossip can have the pro-social function of protecting others from a potential exploitation [@Feinberg2012]. However, gossip has also some self-interest effects: for instance, observing another subject being exploited and having the opportunity to gossip about the exploiter significantly reduced negative affects in @Feinberg2012 's participants. Moreover, providing correct information while gossiping helps to build a positive reputation [@Giardini2021].

In conclusion, when people have the possibility to communicate before, during or after a cooperative interaction, the cooperation level increases. Even if the strength of this effect can vary depending on the situation, communication never seems to have a negative impact on cooperation.

## Social Norms

Social norms can provide a guide on how to behave, especially when there is uncertainty about others' behaviour [@VonBorgstede2018]. @Thogersen2008 studied the effect of social norms on daily social dilemma situations such as environmental behaviours, and found that both injunctive (what others approve) and descriptive (what others do) norms are positively related to behaviours. Moreover, cooperation can be maintained via social norms, because people who do not respect them are susceptible of social sanctions. In fact, if social approval is possible, people cooperate more [@Rege2004].

There are several explanations of social norms' effect on cooperation. First, according to the appropriateness framework, when taking a decision people ask themselves: "What does a person like me do in a situation like this?" [@Arora2012]. Therefore they need to define the situation as a cooperative task and to strongly identify with the group [@VanLange2013]. This leads to a greater accessibility of group's norms and consequently (if the norms are cooperative) to more cooperation. Testing this assumption, @Arora2012 showed that a greater identification with social group lead to higher levels of cooperation. The second explanation concerns conformity: in general people adapt their cooperation to others' behaviour, but this is usually explained by the reciprocity principle (cooperation enhances the probability that others will do the same, and therefore that the final payoff will be higher) [@Bardsley2005a]. However, these authors demonstrated that people conform to others' behaviour even if this has no impact on their personal payoff. Therefore, people have a general tendency to imitate others to follow social norms.

Finally, social norms can be modified by participants. For instance, @Weber2008 demonstrated that in several social dilemma experiment there are consistent cooperators, i.e. people who always cooperate. The presence of these people can increase other members' cooperation, and this effect is mediated by the perception of the social norms. Moreover, social norms can also be shaped via communication [@Bicchieri2002].

In conclusion, social norms have a significant effect on people's cooperation in social dilemma situations. Since they are quite easily modified [e.g. @Weber2008], shaping cooperative norms can be a powerful instrument to improve cooperation.

## Anonymity

There are not many studies on the influence of anonymity on cooperation, but, according to @Wang2017, most of the factors that can increase cooperation rely on a reduction of anonymity. For instance, it is not possible to punish defectors without knowing who they are. Therefore, in general it seems that non-anonymous situations can significantly increase cooperation and reduce defection [@Wang2017]. This can be explained by the fact that, when their actions are visible, people worry about others' approval [@Rege2004] and that, in non-anonymous situations, contributions can be a self-presentation strategy, because being perceived as pro-social can increase one's social status [@VanVugt2010a]. There are therefore more social incentives to cooperate in non-anonymous, rather than anonymous, situations.

Moreover, this effect is displayed especially in situations when individuals face a cost when cooperating and know that they will not get any personal benefit. For instance, @Butz2021 demonstrated that, in a public good game where a part of contributions was donated to a charity association, disclosing identities increased cooperation, but only if the donation was deducted from participants' payoff. In the same direction, @VanVugt2010a found that, when actions were public, contributions were higher if the public good was already attained or unattainable, and therefore if participants did not have a personal interest in cooperating.

Reducing anonymity in cooperative interactions can be very effective to encourage people to contribute more. A possible way to do this is to reduce group's size. In fact, according to @Romano2016, larger groups lead to de-individuation, because it is more difficult to understand who is defecting. Therefore, it is easier for members to feel anonymous and to perceive the social punishments associated with defection as less probable. Moreover, communication is easier in smaller groups [@VonBorgstede2018]. This assumption is confirmed experimentally by @Jiang2021, who showed than in smaller groups people contributed more often to a public good.

Another effective method to reduce anonymity is to make people publicly promise to cooperate before the cooperative interaction. For instance, in a laboratory experiment modeling a common pool resource, people who publicly committed exploited significantly less the resource than people who committed in private, even if after committing they were not restricted in their harvesting possibilities [@Mosler1993]. This results was also confirmed by @Przepiorka2020, who showed that non-binding promises increased cooperation. Moreover, even a compulsory commitment has a beneficial effect on cooperation, even if the most efficient groups are those where all participants signed voluntary the pledge [@Koessler2021a].

In conclusion, reducing anonymity during interactions is essential to improve cooperation, because this can facilitate the implementation of others cooperation-enhancing techniques [@Wang2017]. Moreover, there are several low-cost method to make situations non-anonymous.

# 

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
